---
title: "Cheat Sheet"
author: "Ananda Leia"
date: "2023-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load once per session:
```{r}
library(tidyverse)
library(janitor)
library(here)
library(naniar)
library(RColorBrewer)
library(paletteer)
```

##Check current working directory
```{r}
getwd()
```

##Cancel use of Scientific Notation
```{r}
options(scipen=999)
```

##Calculations in RStudio
```{r}
mean()
median()
mode()
sd()
```
##Creating Objects
```{r}
x <- 4 #assign a value to a variable
y <- 5 
z <- sum(x, y) #creating a new object from other objects
vector <- c(x, y ,z) #create a vector by concatinating objects
```
##Creating Vectors
```{r}
numeric_vector <- c(1, 2, 3)
character_vector <- c("one", "two", "three")
vector_sequence <- c(1:3)
```
Viewing objects within a vector
```{r}
vector == 15 #pulls out objects with a certain numeric value
vector == "character" #objectes with a certain character value
vector[3] #object in the third position of the vector
# use <, >, <=, >=, !=
```
## Create Matrices
```{r}
combined_vector <- c(vector1, vector2, vector3)
matrix_name <- matric(combined_vector, nrow = 3, byrow = T)
```
Naming columns and rows of matrix, colnames() or rownames()
```{r}
row_vector <- c(name1, name2)
rownames(matrix_name) <- row_vector
```
Creating new column, cbind() or rbind()
```{r}
new_col <- rowSum(old_matrix)
new_matrix <- cbind(old_matrix, new_col)
```

## Types of Data
Five frequently used `classes` of data: numeric, integer, character, logical, complex.
```{r}
my_numeric <- 42
my_integer <- 2L #adding an L automatically denotes an integer
my_character <- "universe"
my_logical <- FALSE
my_complex <- 2+4i
```
```{r}
class(object) #find out what kind of data an object is
is.numberic(object) #find out if a specific object is a specific data class
as.numberic(object) #convert object to specific data class
```

##Built in Datasets
```{r}
library("gapminder") #country dataset
```

##Dealing with NA's
```{r}
is.na(object)
any.na(object)
```
Remove NA's
```{r}
mean(vector, na.rm = T) #removes NA's from given calculation
```
Convert symbols into NA's
```{r}
"dataset_name" <- read_csv(("file/dataset.csv"), na = c("", "NA", "\\", "-999"))
```

